---
title: "Geospatial Data Analysis"
output: html_notebook
---



```{r}
library(sf)
my_sf = st_read("india_states_2014/india_states.shp")

```


```{r}
library(GADMTools)
india_wrapper = gadm_sf.loadCountries("IND", level = 1, basefile = "./")
```

```{r}
my_spdf = as(my_sf, "Spatial")
class(my_spdf)
```

```{r}
str(my_spdf, max.level = 2)
```

```{r}
library(tidyverse)
glimpse(my_spdf@data)
```

```{r}
ind_sf = st_as_sf(my_spdf)
class(ind_sf)
```
We notice that our "sp" was a SpatialPolygonDataFrame, but now we see a shift towards something consisting of two simultaneous classes: sf and data.frame. Let us print the first observations and see the difference. 

```{r}
head(ind_sf, 4)
```

Personally, I am more happy to analyse data in this format.
The __epsg__ and __proj4string__ determine the co-ordinate reference system (CRS), this is a geographic CRS, giving the longitude and latitude of the states.

```{r}
glimpse(ind_sf)
```

Spatial dataframes in __sf__ are __dataframes__ and can be manipulated as such by means of packages such as __dplyr__. Let's have a go at changing some of the column names like we would with any traditional __dataframe__. Currently the data consists of many names of "states", but India has 29 states and 7 union territories, this in not yet evident from the above data. Let's then manipulate the data to split the __names__ into __states__ and __union territories__. 

```{r}
ind_sf$name
length(ind_sf$name)
```


```{r}
library(dplyr)
uts = c("Delhi", "Andhra Pradesh", "Daman and Diu", "Dadra and Nagar Haveli", "Puducherry", "Andaman and Nicobar Islands", "Lakshadweep", "Chandigarh")
ind_sf <- ind_sf %>% 
  select(name, abbr) %>%
  mutate(type = ifelse(name %in% uts, "Union Territory", "State")) %>%
  rename(abb = abbr, state_ut = name)
  
```


```{r}
glimpse(ind_sf)
```

Note that the __geometry__ variable is __sticky__ and will not drop unless explicitly done so with ind_sf %>% st_set_geometry(NULL).

```{r}
attributes_df <- readRDS("attributes.rds")

ind_sf <- ind_sf %>%
  left_join(attributes_df, by = "state_ut") %>%
  mutate(
    per_capita_gdp_inr <- nominal_gdp_inr / pop_2011,
    per_capita_gdp_usd <- nominal_gdp_usd / pop_2011
  )
```
The attribute data has now been joined and the addition of per capita GDP has been joined to the dataset. Let's jump into some area calculations with the implementation of the __units__ library. 
```{r}
library(units)
library(lwgeom)
# Mutate the area
ind_sf <- ind_sf %>%
  mutate(my_area = st_area(.))

# Convert the units
units(ind_sf$my_area) <- with(ud_units, km^2)

# Mutate GDP density 
ind_sf <- ind_sf %>%
  mutate(gdp_density_usd_km2 = nominal_gdp_usd / my_area)

```
We have now used units to successfully add columns utilizing units such as km^2, let's take it a step further and start having a look at the geometry of our data. 

```{r}
ind_sf <- ind_sf %>%
  mutate(my_area = as.vector(my_area), 
         gdp_density_usd_km2 = as.vector(gdp_density_usd_km2))

original_geometry <- st_geometry(ind_sf)

library(rmapshaper)
simp_sf <- ms_simplify(ind_sf, keep = 0.01, keep_shapes = TRUE)
simple_geometry <- st_geometry(simp_sf)

par(mfrow = c(1,2))
plot(original_geometry, main = "Original Geometry")
plot(simple_geometry, main = "Simple Geometry")

```

It's very clear when using __ms_simplify__ that the image is much smoother and pleasing to work with. This is due to the fact that __ms_simplify__ only uses 1% of the vertices, whilst still maintaining the geometry of the shapes, making for a smoother image. 
Another significant benefit is that the storage of the Simple Geometry plot is 150KB compared to the 9.56MB Original plot.

```{r}
library(pryr)
object.size(original_geometry)
```

```{r}
object.size((simple_geometry))
```

```{r}
saveRDS(simp_sf, "simp_sf.rds")
```

# Thoughts thus far

It has been really eye-opening to see the various applications of working with geospatial data. The main take away has been the understanding of the way in which processing geospatial and attribute data is done with the help of a __tidy__ work flow. 
The next step, static mapping, will certainly be an interesting development and I very much look forward to delving deeper into the workings behind it. 


# Creating Static Maps in R

This sections aims to illustrate the visual power associated with translating data into visual representations of static maps. Although there are a plethora of incredible packages available to achieve successful results, the main package of interest for this section is __ggplot2__, however let's start with some of the more simple plots and start to realise the appropriateness of a specific plot for a specific objective.


```{r}
library(tidyverse)
library(sf)
simp_sf <- readRDS("simp_sf.rds")
plot(simp_sf["pop_2011"])
```

## Thematic Maps (with tmaps)

Let's experiment with tmaps to create a visualisation which filters out Union Territories as well as provides a categorical representation of the population.

```{r}
library(tmap)
simp_sf %>%
  filter(!state_ut %in% c("Andaman & Nicobar Islands", "Lakshadweep"
)) %>%
  tm_shape() +
  tm_fill(col = "pop_2011", title = "No. People") + 
  tm_borders(lwd = 0.5) +
  tm_text("abb", size = 0.5) +
  tm_style("gray") +
  tm_layout(
    main.title = "Population (2011)",
    main.title.position = c("center"),
    main.title.size = 1,
    legend.position = c("right", "bottom")
  ) +
  tm_credits("Data:\n2011 Census", position = c("left", "bottom"))

```

```{r}
library(tmap)
states_sf <- simp_sf %>%
  filter(!type == "Union Territory")

growth <- tm_shape(states_sf) +
  tm_fill(col = "decadal_growth", title = "Percentage") + 
  tm_borders(lwd = 0.5) +
  tm_layout(
    main.title = "Population Growth of States (2001-2011)",
    main.title.position = c("center"),
    main.title.size = 1,
    legend.position = c("right", "bottom")
  ) +
  tm_credits("Data:\n2001-2011 Census", position = c("left", "bottom"))
  
density <- tm_shape(states_sf) +
  tm_fill(col = "density_km2", title = "No. People / Sq Km", palette = "YlGnBu") + 
  tm_borders(lwd = 0.5) +
  tm_layout(
    main.title = "Population Density (2011)",
    main.title.position = c("center"),
    main.title.size = 1,
    legend.position = c("right", "bottom")
  ) +
  tm_credits("Data:\n2011 Census", position = c("left", "bottom"))

tmap_arrange(growth, density)

```
Another very useful aspect of tmaps is the ability to plot inset maps; plotting only part of a map and analysis data from just one part of the whole. Let's investigate the sex ratio in Northeastern India. 

```{r}
ne_sex <- simp_sf %>%
  filter(region == "Northeastern")

growth <- tm_shape(states_sf) +
  tm_fill(col = "sex_ratio", title = "Sex Ratio", palette = "-Reds") + 
  tm_borders(lwd = 0.5) +
  tm_layout(
    main.title = "Sex Ratio (2011)",
    main.title.position = c("center"),
    main.title.size = 1
  ) +
  tm_credits("Data Source: Wikipedia", position = c("left", "top"))

regional <- simp_sf %>%
  group_by(region) %>%
  summarise(pop = sum(pop_2011))

inset <- regional %>%
  filter(!region == "Arabian Sea",
         !region == "Bay of Bengal") %>%
  mutate(northeast = ifelse(region == "Northeastern", TRUE, FALSE)) %>%
  tm_shape() +
  tm_fill(col = "northeast", palette = c("gray", "red")) +
  tm_style("cobalt") +
  tm_legend(show = FALSE)

library(grid)
ne_sex
print(inset, vp = viewport(0.24, 0.18, width = 0.2, height = 0.4))




```


When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file).
